---
title: "Figure 5:  gene pathway; triple mutants; TMB adjusted vs multi-cancer model"
site: workflowr::wflow_site
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
editor_options:
  chunk_output_type: console
---

```{r caching, echo=FALSE}
knitr::opts_chunk$set(autodep = TRUE)
```

# Panel A: Gene pathway results


```{r packages, message=FALSE, warning=FALSE}
library(ggplot2)
library(plyr)
library(reshape2)
library(cowplot)
library(tidyverse)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(readxl)
library(scales)
library(here)
library(ggrepel)
library(purrr)
library(rasfuns)
```

```{r 5bdata, message=FALSE, warning=FALSE}
here <- here::here
# Read any mutations csv for gene pathways
comut_gp <- here("public",
                 "table",
                 "co_mutation_allele.Rmd",
                 "any_mutation_(gene_pathways).csv") %>%
    read_csv(show_col_types=FALSE)
# list of co-muts and cancer types to plot
plot_gp <- read_excel( here( "data","ras_pathway_pval.xlsx"))
comut_gp <-  comut_gp %>%
    mutate(ras = gsub("_1", "_G1",ras ))%>%
    mutate(ras = gsub("_6", "_Q6",ras ))%>%
    mutate(index = paste0(cancer, "~",
                          ras, "~", nonras))
gene_pathway.levels <- table(plot_gp %>% select(gene_pathway)) %>%
    as.data.frame() %>%
    arrange(desc(Freq))
gene_pathway.levels <- as.character(gene_pathway.levels$Var1)
plot.df <- plot_gp %>%
    mutate(ras = gsub(":", "_",ras ))%>%
    mutate(index = paste0(cancer_full, "~", ras, "~", gene_pathway))%>%
    merge(., comut_gp %>% select(index, p, `50%`), by.x = "index", by.y = "index" , all.x = TRUE)%>%
    mutate(cancer = factor(cancer, levels = unique(cancer))) %>%
    mutate(gene_pathway = factor(gene_pathway, levels = gene_pathway.levels)) %>%
    mutate(ras = factor(ras, levels = unique(ras))) %>% mutate(y=0)%>%
    mutate(ras_cancer = paste0(cancer, '~', ras))%>%
    mutate(ras_cancer = factor(ras_cancer, levels = unique(ras_cancer))) %>%
    mutate(comut_type = ifelse(`50%`>0, "co-occurence" , "mutual-exclusivity"))%>%
    mutate(comut_type =factor(comut_type, levels = unique(comut_type)) )%>%
    select(-c("logp", "type"))
```

```{r fig5a, message=FALSE, warning=FALSE, fig.height = 12, fig.width=10, dev=c("png", "pdf")}
col.pal <- c("#0066cc","#cc0000")
names(col.pal) <- c( 'Mutual\nexclusivity\n', '\nCo-occurence\n')
dat.list <- plot.df %>%
    mutate(gene_pathway=case_when(gene_pathway=="RAS_RAF_MAPK_brief"~"RAS_RAF_MAPK",
                                  gene_pathway=="RAS_RAF_MAPK_extended"~"RAS_RAF_MAPK",
                                  TRUE~as.character(gene_pathway)),
           comut_type=case_when(comut_type=="mutual-exclusivity"~"Mutual\nexclusivity\n",
                                comut_type=="co-occurrence"~"\nCo-occurrence\n")) %>%
    group_by(gene_pathway) %>%
    nest()
pathways <- tibble(label=c("RAS/RAF/MAPK pathway",
                           "NRF2 pathway/Oxidative stress response",
                           "PI3K/AKT pathway",
                           "Cell cycle progression",
                           "Chromatin regulation",
                           "DNA damage response"),
                   gene_pathway=dat.list$gene_pathway[c(3, 2, 1, 6, 4, 5)]) %>%
    mutate(label=factor(label, label))
dat.list2  <- dat.list %>%
    left_join(pathways, by="gene_pathway") %>%
    arrange(label)
lollipop_plot <- function(dat){
    dat %>%
        ggplot(aes(ras_cancer, p,
                   fill = comut_type,
                   color = comut_type)) +
        geom_point(aes(fill=comut_type))+
        geom_linerange(aes(ymin = y, ymax = p,
                           color = comut_type))+
        coord_flip()+
        scale_fill_manual(values = col.pal, name = '',
                          guide = guide_legend(position = 'none',
                                               direction = "vertical")) +
        scale_color_manual(values = col.pal,
                           name = '',
                           guide = guide_legend(position = 'none',
                                                direction = "vertical")) +
        ylab("-log10 p")+
        xlab("")+
        theme_classic(base_size=15) +
        theme(axis.line=element_blank(),
              axis.ticks.x = element_blank(),
              plot.margin = margin(0,0,0,0, "cm"),
              axis.text.y = element_text(size=13, angle = 0),
              ##axis.text.x = element_text(size=10, angle =0),
              axis.text.x=element_blank(),
              axis.title.x=element_blank(),
              panel.grid= element_blank() ,
              panel.border = element_rect(fill=NA),
              plot.title = element_text(hjust=0.5, size = 10)) +
        ylim(c(0, 9)) +
        guides(color="none", fill="none")
}
gl <- map(dat.list2$data, lollipop_plot)
names(gl) <- dat.list2$label
for(i in 1:6){
    gl[[i]] <- gl[[i]] + ggtitle(names(gl)[i]) +
        theme(plot.title=element_text(size=16, hjust=0),
              plot.margin=unit(c(1.5, 1.5, 0, 1.5), "lines"))
}
gl[[6]]  <- gl[[6]] +
    theme(axis.text.x=element_text(size=13),
          axis.title.x=element_text(size=17)) +
    ylab(expression(-log[10](p)))
tmp <- gl[[6]] +
    theme(legend.text=element_text(size=17)) +
    guides(color=guide_legend(title=""))
leg <- cowplot::get_legend(tmp)
gl <- map(gl, ggplotGrob)
w <- gl[[1]]$widths
for(i in 2:6) gl[[i]]$widths <- w
gl[[7]] <- leg
heights <- map_int(dat.list2$data, nrow)
heights[heights==1] <- 1.5
## compensate for x-axis and title
heights[6] <- 2.2
grid.arrange(grobs=gl,
             widths=c(1, 0.22),
             heights=heights,
             layout_matrix=cbind(1:6, 7))
A  <- arrangeGrob(grobs=gl,
                  widths=c(1, 0.22),
                  heights=heights,
                  layout_matrix=cbind(1:6, 7))
## todo
## add title to each panel
## use classic theme
```


# Panel B: Triple mutations

```{r data, message=FALSE, warning=FALSE}
load(here("output", "co-mutation","kras12c_tripleMuts.rda"))
threshold <- 1
plot.df <- c()
for(cancer in names(kras12c_tripleMuts)){
    tmp.df <- kras12c_tripleMuts [[cancer]] %>%
        filter(intersect.percent >= threshold)%>%
        mutate(cancer_type = cancer)
    plot.df = rbind(plot.df ,tmp.df )
}
plot.df  <-  plot.df %>%
    mutate(cancer_type = recode(cancer_type, "non-small cell lung cancer"="NSCLC" ,
                                "cancer of unknown primary"  = "CUP",
                                "colorectal cancer" = "CRC",
                                "pancreatic cancer" = "PAC",
                                "uterine cancer" = "Uterine")) %>%
    mutate(gene = paste0("KRAS_", gene)) %>%
    arrange(desc(intersect.percent))
cancer.levels <- c("CUP", "CRC" ,
                   "PAC", "Uterine",
                   "NSCLC")
plot.df$cancer_type <- factor(plot.df$cancer_type, levels = cancer.levels)

gene.levels <- data.frame(table(plot.df$gene)) %>% arrange(desc(Freq))
gene.levels <- gene.levels$Var1
plot.df <- plot.df %>%
    arrange(desc(intersect.percent))  %>%
    arrange(factor(gene, levels = gene.levels))
tmp.df <- plot.df[-c(1:4), ] %>%
    arrange(desc(intersect.percent))  %>%
    arrange(factor(cancer_type, levels = cancer.levels))
plot.df <- rbind(plot.df[c(1:4), ], tmp.df)
```



```{r fig5b, message=FALSE, warning=FALSE,  fig.height = 4, fig.width = 8, dev=c("png", "pdf")}
base <- expand.grid(unique(plot.df$cancer_type),
                    unique(plot.df$gene))
base$gene <- rev(base$gene)
colnames(base) <- c('cancer_type', 'gene')
base$cancer_type <- factor(base$cancer_type,
                           levels = unique(cancer.levels))
base$gene <- factor(base$gene,
                    levels = rev(unique(base$gene)))
genelevels <- unique(base$gene[1:100]) %>%
    str_replace_all("_", " / ") %>%
    str_replace_all("KRAS / ", "")
base2 <- base[c(1:100), ] %>%
    mutate(gene=as.character(gene)) %>%
    mutate(gene=str_replace_all(gene, "_", " / ")) %>%
    mutate(gene=str_replace_all(gene, "KRAS / ", "")) %>%
    mutate(gene=factor(gene, genelevels))
plot.df <- plot.df %>%
    mutate(gene=as.character(gene)) %>%
    mutate(gene=str_replace_all(gene, "_", " / ")) %>%
    mutate(gene=str_replace_all(gene, "KRAS / ", "")) %>%
    mutate(gene=factor(gene, genelevels))
cm4 <- ggplot(base2, aes(x = gene, y = cancer_type)) +
    geom_tile(fill = '#ffffff',
              color = '#deebf7',
              width = 0.8,
              height = 0.8) +
    theme_minimal() +
    geom_tile(data = plot.df %>%
                  subset(gene %in% unique(base2[ , 'gene'])),
              aes(fill = intersect.percent),
              size = 0.8,
              width = 0.8,
              height= 0.8) +
    scale_fill_distiller(palette = 'Reds',
                         name = 'Prevalence (%)',
                         values = rescale(1:10),
                         guide = guide_colorbar(direction = "vertical"),
                         trans = "reverse") +
    theme(#axis.ticks.x = element_blank(),
        plot.margin = margin(-0.5,-0.5,-0.5,-0.5, "cm"),
        axis.text.x = element_text(size = 12,
                                   angle = 70,
                                   hjust=1),
        axis.text.y = element_text(size = 15, angle = 0),
        panel.grid = element_blank(),
        title = element_text(size = 8),
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.title=element_text(size=16),
        legend.text=element_text(size=14),
        plot.title=element_text(size=18))+
    ylab("") + xlab("") +
    ggtitle("KRAS G12C triple mutants")
aggregate.plot3 <- plot_grid(NULL, cm4, NULL ,
                             nrow = 1,
                             rel_widths = c(0.25, 8, 0.25))
B <- plot_grid(NULL, aggregate.plot3,
               NULL , ncol = 1,
               rel_heights = c(0.30, 8, 0.25))
Bg <- ggplotGrob(B)
B
```

# Panel C: Concordance of estimated gene-gene dependencies between TMB adjusted model and multi-cancer model

```{r figoptions, echo=FALSE}
knitr::opts_chunk$set(autodep = TRUE, warning=FALSE)
base.size <- 16
gene.size <- 3
point.size <- 3
strip.size <- 19
axis.text <- 16
axis.title <- 17
n.size <- 4
bar.height <- 0.2
letter.size <- 2.5 ## size of letters for caption labels
```


```{r 5c_data}
#source(here("code", "co-mutation", "stan_funs.R"))
##devtools::load_all(here("code", "rasfuns"))
cancerlabs <- read_csv(here("data", "cancer_labels.csv"))
colors <- readRDS(here("output", "co-mutation",
                       "summarize_models.R",
                       "colors.rds")) %>%
    left_join(cancerlabs, by="cancer") %>%
    mutate(cancer=cancer_label)
cols <- setNames(colors$color, colors$cancer)
```

```{r tmb_labels}
extdir <- system.file("extdata", package="genie.6.1")
fname <- file.path(extdir, "tumor_normal_predictions.csv")
data(mutation_tmb, package="contingency.table")
data(patient_universe, package="genie.6.1")
matched_universe <- filter(patient_universe, is_matched)
tmb_labels <- read_csv(fname) %>%
    rasfuns:::attach_tmb_label(matched_universe)
```

Select the following:
STK11 and KEAP1 in NSCLC
TP53, inactivating APC in CRC

```{r read_mutations}
comutation <- here("output", "co-mutation", "summarize_models.R",
                   "combined.rds") %>%
    readRDS() %>%
    unnest(data) %>%
    ungroup() %>%
    left_join(cancerlabs, by="cancer") %>%
    mutate(cancer=cancer_label) %>%
    group_by(cancer_group, alteration_group, confounder) %>%
    nest()

## significant associations in multi-cancer model
## -log10p > 6
sig_multicancer <- rasfuns:::significant_multicancer
filter_sig <- rasfuns:::filter_by_signif
signif <- sig_multicancer(comutation,
                          group_="Any mutation",
                          ras_="RAS_12-61",
                          p_cutoff=6)
isignif <- sig_multicancer(comutation,
                           group_="Inactivating mutation",
                           ras_="RAS_12-61",
                           p_cutoff=6) %>%
    filter(!nonras %in% signif$nonras & cancer %in% signif$cancer)
signif <- bind_rows(signif, isignif)
add_tmb_label <- rasfuns:::add_tmb_label
filter_by <- rasfuns:::filter_by
x  <- comutation %>%
    filter_by(c("Any mutation",
                "Inactivating mutation"),
              "RAS_12-61", "TMB") %>%
    add_tmb_label(tmb_labels) %>%
    filter_sig(signif)
ylist <- rasfuns:::list_signif_by(signif, x$uid)
effect_list <- rasfuns:::list_confounding(ylist, x)
```

## Co-mutation of RAS (any codon) and any non-RAS mutation

Identify co-mutations in the stratified analysis with 95% credible intervals that do not overlap the 95% credible interval in the multi-cancer model.  Focus on the set of co-mutations in which the association in the multi-cancer model was statistically significant.

Several co-mutations significant in the multi-cancer model are not in the stratified analyses.  In the stratified analyses, we required that the number of RAS mutations and the number of non-RAS mutations must each be > 100 (l. 133, `contingency.table/data-derived/mutations_tmb.R`).

Below, we plot overall versus multi-cancer, highlighting the genes with apparent discordance in 1 or more TMB strata.

```{r concordance, fig.width=12, fig.height=7, dev=c("png", "pdf")}
rename <- dplyr::rename
multicancer <- effect_list$data[[1]] %>%
    select(nonras, alteration_group, cancer, `50%`) %>%
    unite("uid", c(nonras, alteration_group, cancer)) %>%
    rename(x="50%")
tmb  <- effect_list$data[[2]] %>%
    map_dfr(function(x) filter(x, grepl("mu_beta", Parameter))) %>%
    select(nonras, alteration_group, cancer, `50%`, `2.5%`, `97.5%`) %>%
    unite("uid", c(nonras, alteration_group, cancer)) %>%
    rename(y="50%")
dat <- left_join(multicancer, tmb, by="uid") %>%
    separate(uid, c("nonras", "alt_group", "cancer"), sep="_")
multicancer2 <- effect_list$data[[3]] %>%
    select(nonras, alteration_group, cancer, `50%`) %>%
    unite("uid", c(nonras, alteration_group, cancer)) %>%
    rename(x="50%")
tmb2  <- effect_list$data[[4]] %>%
    map_dfr(function(x) filter(x, grepl("mu_beta", Parameter))) %>%
    select(nonras, alteration_group, cancer, `50%`, `25%`, `75%`) %>%
    unite("uid", c(nonras, alteration_group, cancer)) %>%
    rename(y="50%")
dat2 <- left_join(multicancer2, tmb2, by="uid")  %>%
    separate(uid, c("nonras", "alt_group", "cancer"), sep="_")
all <- bind_rows(dat, dat2) %>%
    mutate(cancer=factor(cancer, sort(unique(cancer))))
dat <- dat %>%
    mutate(cancer=factor(cancer, levels(all$cancer)))
##stop()
cols <- cols[levels(all$cancer)]
C <- all %>%
    ggplot(aes(x, y)) +
    geom_abline(intercept=0, slope=1) +
    geom_hline(yintercept=0, linetype="dashed", color="gray") +
    geom_vline(xintercept=0, linetype="dashed", color="gray") +
    geom_errorbar(aes(ymin=`25%`, ymax=`75%`),
                  color="gray", width=0.01) +
    geom_point(aes(color=cancer), size=point.size) +
    theme_bw(base_size=base.size) +
    ##geom_point(size=point.size, pch=21) +
    geom_point(data=dat,
               aes(color=cancer)) +
    geom_label_repel(data=dat,
                    aes(label=nonras),
                    fill="white", segment.color="black",
                    ##alpha=0.3,
                    size=gene.size,
                    max.overlaps=Inf,
                    box.padding=0.5) + ##color="transparent") +
    ylab("Cross-TMB average") +
    xlab("Association from multi-cancer model") +
    theme(panel.grid=element_blank(),
          legend.position=c(0.1, 0.8),
          ##legend.position="bottom",
          strip.background=element_blank(),
          strip.text=element_text(size=strip.size),
          axis.text=element_text(size=axis.text),
          axis.title=element_text(size=axis.title)) +
    xlim(c(-5, 3)) + ylim(c(-5, 3)) +
    guides(color=guide_legend(title="")) +
    scale_color_manual(values=cols)
##A_leg <- cowplot::get_legend(fig)
##A <- fig + guides(color=FALSE)
C
Cg <- ggplotGrob(C)
```

# Panel D: Gene-gene dependencies within TMB strata

### plot intervals for overall and multicancer for genes with possible tmb confounding


```{r paneld}
goi <- c("STK11", "KEAP1", "TP53", "APC")
multicancer <- effect_list$data[[1]] %>%
    select(nonras, alteration_group,
           cancer, n, `50%`, `2.5%`, `97.5%`) %>%
    unite("uid", c(nonras, alteration_group, cancer), remove=FALSE) %>%
    mutate(analysis="Multi-cancer") %>%
    filter(nonras %in% goi)
tmb  <- effect_list$data[[2]] %>%
    map_dfr(function(x) filter(x, grepl("mu_beta", Parameter))) %>%
    select(nonras, alteration_group,
           cancer, n, `50%`, `2.5%`, `97.5%`) %>%
    unite("uid", c(nonras, alteration_group, cancer), remove=FALSE) %>%
    filter(uid %in% multicancer$uid) %>%
    mutate(analysis="Overall\n(TMB model)")
dat <- bind_rows(multicancer, tmb) %>%
    select(-uid) %>%
    unite(uid, c("nonras", "cancer"), sep=" in ", remove=FALSE)
dat <-  dat %>%
    filter(!uid %in% "TP53 in Pancreatic") %>%
    mutate(analysis=factor(analysis,
                           c("Multi-cancer", "Overall\n(TMB model)")),
           analysis=fct_rev(analysis))
tmb  <- effect_list$data[[2]] %>%
    map_dfr(function(x) filter(x, grepl("^beta", Parameter))) %>%
    select(nonras, alteration_group,
           cancer, label, n, `50%`, `2.5%`, `97.5%`) %>%
    unite("uid", c(nonras, alteration_group, cancer), remove=FALSE) %>%
    filter(uid %in% multicancer$uid) %>%
    select(-uid) %>%
    unite("uid", c(nonras, cancer), sep=" in ", remove=FALSE)
levels <- c(levels(dat$analysis), levels(tmb$label))
dat2 <- dat %>%
    rename(label=analysis)  %>%
    bind_rows(tmb) %>%
    mutate(label=factor(label, levels))

dat3 <- dat2 %>%
    filter(!uid %in% "TP53 in Pancreatic") %>%
    mutate(uid=factor(uid,
                      levels=c("KEAP1 in NSCLC",
                               "STK11 in NSCLC", "TP53 in NSCLC",
                               "APC in CRC", "TP53 in CRC")))
D <- dat3 %>%
    ggplot(aes(`50%`, label)) +
    geom_errorbarh(aes(xmin=`2.5%`, xmax=`97.5%`,
                       color=cancer), height=bar.height,
                   size=1.5) +
    geom_point(##pch=21, fill="white",
               size=point.size, aes(color=cancer)) +
    theme_bw(base_size=base.size) +
    geom_text(aes(x=Inf, y=label, label=n), hjust=1,
              size=n.size, color="gray30") +
    ylab("TMB strata") +
    xlab("Log odds ratio") +
    geom_hline(yintercept=2.5) +
    geom_vline(xintercept=0, linetype="dashed", color="gray") +
    theme(panel.border=element_rect(color="gray"),
          axis.title=element_text(size=axis.title+1),
          ##axis.title.y=element_blank(),
          strip.text.y=element_blank(),
          strip.text.x=element_text(size=axis.title+2),
          axis.text.x=element_text(size=axis.text),
          axis.text.y=element_text(size=axis.text),
          panel.grid=element_blank(),
          legend.position="bottom",
          strip.background=element_blank()) +
    facet_wrap(~uid, ncol=5) +#, nrow=1) +
    scale_color_manual(values=cols) +
    guides(color='none')
D
Dg <- ggplotGrob(D)
```

# Figure 5

```{r fig5, fig.width=25, fig.height=17, dev=c("png", "pdf")}
grid.newpage()
gl <- list(A,
           Bg,
           Cg,
           Dg,
           grob(),
           grob(),
           grob())
vp <- viewport(width=unit(0.98, "npc"),
               height=unit(0.98, "npc"))
pushViewport(vp)
grid.arrange(grobs=gl,
             heights=c(1, 0.05, 1, 0.1, 1),
             widths=c(1, 0.1, 1),
             newpage=FALSE,
             layout_matrix=rbind(c(1, 6, 2),
                                 c(1, 6, 7), ## space between panels B and C
                                 c(1, 6, 3),
                                 c(5, 6, 5),
                                 c(4, 6, 4)))
upViewport()
grid.text("A", x=unit(0.015, "npc"), y=unit(0.98, "npc"),
          gp=gpar(cex=2.5))
grid.text("B", x=unit(0.5, "npc"), y=unit(0.98, "npc"),
          gp=gpar(cex=2.5))
grid.text("C", x=unit(0.5, "npc"), y=unit(0.67, "npc"),
          gp=gpar(cex=2.5))
grid.text("D", x=unit(0.015, "npc"), y=unit(0.33, "npc"),
          gp=gpar(cex=2.5))
gc()
```
