---
title: "Figure S15: co-occurrence and mutation signatures"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---


```{r caching, echo=FALSE}
knitr::opts_chunk$set(autodep = TRUE, warning=FALSE, echo=FALSE)
base.size <- 16
gene.size <- 3
point.size <- 3
strip.size <- 19
axis.text <- 14
axis.title <- 17
n.size <- 4
bar.height <- 0.2
letter.size <- 2.5 ## size of letters for caption labels
```
In particular, RAS codon 12, 13 and 61 mutations co-occurred less frequently with {AMER1}, {KEAP1}, {STK11} and {TP53} in NSCLC tumors harboring a C>A rich smoking related signature (Supplementary Figure \ref{sfig:msig}). In CRC, {KRAS} codon 12, 13 and 61 and {PIK3CA} mutations co-occurred more frequently in tumors that harbored an APOBEC mutational signature (Supplementary Figure \ref{sfig:msig}). Interestingly, in NSCLC and consistent with the TMB stratified analyses, {KRAS} G12C and {KEAP1} or {STK11} co-mutations were less frequent in tumors harboring a smoking-associated mutational signature. {KRAS} G12C and {STK11} co-mutations were less frequent in tumors harboring an APOBEC mutational signature.

# Association between co-occurrence and mutation signatures

Analyses limited to mutations in non-RAS genes (does not include copy number alterations or rearrangements in non-RAS genes).

```{r packages, message=FALSE, warning=FALSE}
library(devtools)
library(magrittr)
library(tidyverse)
library(genie.6.1)
library(kableExtra)
library(grid)
library(purrr)
library(gridExtra)
library(here)
library(ggrepel)
dat.list <- here("output", "mutsig_s13",
                 "data_s13.rds") %>%
    readRDS()
cancerlabs <- here("data", "cancer_labels.csv") %>%
    read_csv(show_col_types=FALSE)
colors <- here("output", "co-mutation",
               "summarize_models.R",
               "colors.rds") %>%
    readRDS() %>%
    left_join(cancerlabs, by="cancer") %>%
    mutate(cancer=cancer_label)
cols <- setNames(colors$color, colors$cancer)
```

```{r A}
panel_a <- dat.list[["A"]]
cols <- cols[sort(names(cols))]
A <- panel_a %>%
    ggplot(aes(x, `50%`)) +
    geom_abline(intercept=0, slope=1) +
    geom_hline(yintercept=0, linetype="dashed", color="gray") +
    geom_vline(xintercept=0, linetype="dashed", color="gray") +
    geom_errorbar(aes(ymin=`25%`, ymax=`75%`),
                  color="gray", width=0.01) +
    geom_point(aes(color=cancer), size=point.size) +
    theme_bw(base_size=base.size) +
    ylab("Mutation signature average") +
    xlab("Association from multi-cancer model") +
    theme(panel.grid=element_blank(),
          strip.background=element_blank(),
          strip.text=element_text(size=strip.size),
          axis.text=element_text(size=axis.text),
          axis.title.x=element_text(size=axis.title),
          axis.title.y=element_text(size=axis.title,
                                    vjust=0),
          legend.position=c(0.15, 0.7)) +
    xlim(c(-5, 3)) + ylim(c(-5, 3)) +
    guides(color=guide_legend(title="")) +
    scale_color_manual(values=cols)
Ag <- ggplotGrob(A)
```

```{r B}
panel_b <- dat.list[["B"]]
B <- panel_b %>%
    ggplot(aes(`50%`, label)) +
    geom_errorbarh(aes(xmin=`2.5%`, xmax=`97.5%`,
                       color=cancer),
                   height=bar.height,
                   size=1.1) +
    geom_point(pch=21, fill="white",
               size=point.size,
               aes(color=cancer)) +
    theme_bw(base_size=base.size) +
    geom_text(aes(x=Inf, y=label, label=n), hjust=1,
              size=n.size, color="gray30") +
    ylab("Mutation signature") +
    xlab("Log odds ratio") +
    geom_hline(yintercept=2.5) +
    geom_vline(xintercept=0, linetype="dashed",
               color="gray") +
    theme(panel.border=element_rect(color="gray"),
          axis.title.y=element_text(hjust=1,
                                    size=axis.title,
                                    color="white"),
          strip.text.y=element_blank(),
          axis.text.x=element_text(size=axis.text),
          axis.text.y=element_text(size=axis.text),
          panel.grid=element_blank(),
          legend.position="bottom",
          strip.background=element_blank()) +
    facet_wrap(~uid) +#, nrow=1) +
    scale_color_manual(values=cols) +
    guides(color="none")
Bg <- ggplotGrob(B)
```

```{r fig_s15, fig.width=12, fig.height=15, dev=c("png", "pdf")}
##index <- grep("panel-1", Bg$layout$name)
##Ag$widths[1:4] <- Bg$widths[1:4]
gl <- list(Ag, Bg, nullGrob())
grid.arrange(grobs=gl,
             widths=c(0.08, 1),
             heights=c(1/3, 2/3),
             layout_matrix=rbind(c(4, 1),
                                 c(2, 2)))
```
