---
title: "Figure S9"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

# RAS and KRAS prevalence stratified by sex


```{r caching, echo=FALSE}
knitr::opts_chunk$set(autodep = TRUE)
```

```{r packsexs, messsex=FALSE, warning=FALSE}
library(here)
library(magrittr)
library(tidyverse)
library(ggplot2)
rename <- dplyr::rename
```

```{r read_data}

multi <-here("public", "table",
            "ras.Rmd",
            "multicancer.csv")%>%
        read_csv(show_col_types=FALSE) %>%
        mutate(hugo_symbol=factor(hugo_symbol,
                              c("RAS", "KRAS",
                                "NRAS"))) %>%
         mutate(multicancer_uid = paste(cancer, hugo_symbol, codon, sep = "~"))%>%
         arrange(hugo_symbol)

sex <- here("public", "table",
            "ras.Rmd",
            "sex.csv") %>%
       read_csv(show_col_types=FALSE) %>%
       mutate(hugo_symbol=factor(hugo_symbol,
                              c("RAS", "KRAS", "NRAS"))) %>%
       mutate(multicancer_uid = paste(cancer, hugo_symbol, codon, sep = "~"))%>%
       arrange(hugo_symbol)

sex <- rbind(multi, sex)
```


RAS mutations highlighted in these panels were selected from an assessment of whether the 95% credible interval from the multi-cancer model overlapped the 95% credible intervals within sex strata.  Combinations with no overlap for either male or female prevalence are plotted here.  See `potential_confounding.Rmd` for additional details.



```{r}
uid <- unique(sex$multicancer_uid)
facet_labels <- function(uid){
    uid2 <- strsplit(uid, "~")
    caps <- Hmisc::capitalize
    cancer <- sapply(uid2, "[", 1) %>%
        str_replace_all("cancer of unknown primary", "CUP") %>%
        str_replace_all("non-small cell lung cancer", "NSCLC") %>%
        str_replace_all("pancreatic cancer", "PAC") %>%
        str_replace_all("ovarian cancer", "OVCA") %>%
        str_replace_all("colorectal cancer", "CRC") %>%
        str_replace_all("b-lymphoblastic leukemia/lymphoma",
                        "B ALL") %>%
       str_replace_all("leukemia" ,"Leukemia") %>%
        caps()
    gene <- sapply(uid2, "[", 2)
    codons <- sapply(uid2, "[", 3)
    flabels <- paste0(cancer, ", ", gene, " (", codons, ")") %>%
        setNames(uid)
    flabels
}

flabels <- facet_labels(uid)

# list of labels to be displayed in fig 
fig.labels <- c( 
   "CRC, KRAS (Codon 12)", 
    "CRC, RAS (Codon 12, 13, or 61)", 
    "Melanoma, RAS (Codon 12, 13, or 61)", 
   "NSCLC, KRAS (Codon 12)",
   "NSCLC, KRAS (Codon G12C)",
       "NSCLC, RAS (Codon 12, 13, or 61)"
)
 flabels <- flabels[which(flabels  %in% fig.labels  )] 
 flabels = flabels[match( fig.labels,flabels )]

 # Only retain labels that are to be displayed in fig S6
 sex <- sex %>%
        filter(multicancer_uid %in% names(flabels)) %>%
        mutate(multicancer_uid = factor(multicancer_uid, levels = names(flabels))) %>%
        filter(! is.na(variable)) %>%
         mutate(variable = recode(variable, "Overall" = "Multi-cancer")) 
                   
```


```{r fig_s9, fig.width=12, fig.height=6, dev=c("png", "pdf")}

sexlabels <-  c("Female", "Male", "Multi-cancer")
sex %>%
    mutate(variable=factor(variable, sexlabels),
           variable=droplevels(variable)) %>%
    ggplot(aes(median, variable)) +
    geom_errorbarh(aes(xmin=q1, xmax=q2), height=0.1) +
    geom_point(pch=21, fill="white") +
    ylab("") +
    geom_text(x=Inf, aes(label=paste0(y, "/", n)),
              hjust=1, size=3) +
    theme_bw(base_size=15) +
    scale_x_continuous(expand=expansion(mult=0.5)) +
    theme(panel.grid=element_blank()) +
    xlab("Prevalence") +
    facet_wrap(~multicancer_uid, scales="free_x",
               labeller=labeller(multicancer_uid=flabels)) +
    theme(strip.background=element_blank()) +
    geom_vline(data=filter(sex, variable=="Multi-cancer") ,
               aes(xintercept=median), linetype="dashed",
               color="gray")
  
```
