---
title: "Patient characteristics associated with RAS prevalence"
site: workflowr::wflow_site
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
editor_options:
  chunk_output_type: console
---


```{r caching, echo=FALSE}
knitr::opts_chunk$set(autodep = TRUE)
```

```{r packages, message=FALSE, warning=FALSE}
library(IRanges)
library(here)
library(magrittr)
library(tidyverse)
library(ggplot2)
library(genie.6.1)
library(grid)
library(gridExtra)
library(kableExtra)
tabdir <- here("public", "table", "potential_confounding.Rmd")
fs::dir_create(tabdir)
library(purrr)
library(viridis)
library(hrbrthemes)
rename <- dplyr::rename
```


# Potential confounding

```{r potential_confounding}
models <- here("output", "prevalence",
               "summarize_models.R",
               "models.rds") %>%
    readRDS()
multicancer <- models %>%
    filter(cancer_group=="multi") %>%
    unnest("data") %>%
    mutate(Y=prettyNum(y, big.mark=","),
           N=prettyNum(n, big.mark=",")) %>%
    unite(nlabel, c(Y, N), sep="/", remove=FALSE) %>%
    group_by(hugo_symbol) %>%
    nest()
models2 <- models %>%
    unnest(data) %>%
    ##filter(ras=="RAS_12-61") %>%
    group_by(cancer) %>%
    nest()

focus <- unnest(multicancer, data) %>%
    filter(y/n > 0.1, n > 100) %>%
    pull(cancer) %>%
    unique()
dat <- models2 %>%
    filter(cancer %in% focus)
dat2 <- unnest(dat, data) %>%
    group_by(cancer, ras) %>%
    nest()
```

Since we do not have the full posteriors saved for evaluating differences from the overall, I've extracted specific strata where the 95% credible interval for RAS prevalence does not overlap the 95% interval from the multi-cancer model.

```{r no_overlap}
no_overlap <- function(cancer, ras, data){
    overall <- filter(data, cancer_group=="multi")
    stratified <- filter(data, cancer_group=="single") %>%
        mutate(ras=ras,
               cancer=cancer)
    r.overall <- IRanges(start=round(overall$q1*1000, 0),
                         end=round(overall$q2*1000, 0))
    r.stratified <- IRanges(start=round(stratified$q1*1000, 0),
                            end=round(stratified$q2*1000, 0))
    is_overlap <- overlapsAny(r.stratified, r.overall)
    keep <- stratified[ !is_overlap , ]
}
results <- dat2 %>%
    pmap(no_overlap) %>%
    map_dfr(bind_rows) %>%
    filter(confounder %in% c("Race", "Sex", "Age")) %>%
    filter(variable != "NA") %>%
    select(-cancer_group) %>%
    select(cancer, ras, hugo_symbol, confounder, variable, y, n, mean,
           median, q1, q2, codon_range, codon)
```

For each row in the above table, pull in all available strata for the potential confounding variable.

```{r}
associations <- results %>%
    group_by(cancer, hugo_symbol, codon, confounder) %>%
    nest() %>%
    unite(uid, c(cancer, hugo_symbol, codon, confounder),
          sep="~", remove=FALSE)
strata_with_association <- dat2 %>%
    unnest(data) %>%
    group_by(cancer, hugo_symbol, codon, confounder) %>%
    nest() %>%
    unite(uid, c(cancer, hugo_symbol, codon, confounder),
          sep="~", remove=FALSE) %>%
    filter(uid %in% associations$uid) %>%
    unite(multicancer_uid, c(cancer, hugo_symbol, codon), sep="~",
          remove=FALSE)
associations2  <- strata_with_association %>%
    unnest("data")
multicancer  <- dat %>%
    unnest(data) %>%
    filter(cancer_group=="multi") %>%
    group_by(cancer, hugo_symbol, codon) %>%
    nest() %>%
    unite(uid, c(cancer, hugo_symbol, codon), remove=FALSE, sep="~") %>%
    filter(uid %in% strata_with_association$multicancer_uid) %>%
    unnest(data) %>%
    ungroup() %>%
    rename(multicancer_uid=uid)

assoc_x_confounder <- group_by(associations2, confounder) %>%
    group_split()
keys <- group_by(associations2, confounder) %>%
    group_keys()
age <- assoc_x_confounder[[which(keys=="Age")]] %>%
    mutate(variable=droplevels(variable)) %>%
    filter(!is.na(variable))
race <- assoc_x_confounder[[which(keys=="Race")]] %>%
    filter(variable != "NA") %>%
    mutate(variable=droplevels(variable)) %>%
    filter(!is.na(variable))
sex <- assoc_x_confounder[[which(keys=="Sex")]] %>%
    mutate(variable=droplevels(variable)) %>%
    filter(!is.na(variable))
```

## Age

```{r age_data}
age_x_hugo <- group_by(age, hugo_symbol) %>%
    group_split()
age_key <- group_by(age, hugo_symbol) %>%
    group_keys()
varlevels <- c("Multi-cancer", levels(age$variable))
mc <- multicancer %>%
    mutate(variable="Multi-cancer",
           variable=factor(variable, varlevels))
ras <- age_x_hugo[[which(age_key=="RAS")]] %>%
    select(-uid) %>%
    mutate(variable=factor(variable, varlevels)) %>%
    bind_rows(filter(mc, multicancer_uid %in% .$multicancer_uid)) %>%
    arrange(multicancer_uid)

kras <- age_x_hugo[[which(age_key=="KRAS")]] %>%
    select(-uid) %>%
    mutate(variable=factor(variable, varlevels)) %>%
    bind_rows(filter(mc, multicancer_uid %in% .$multicancer_uid)) %>%
    arrange(multicancer_uid)

nras <- age_x_hugo[[which(age_key=="NRAS")]] %>%
    select(-uid) %>%
    mutate(variable=factor(variable, varlevels)) %>%
    bind_rows(filter(mc, multicancer_uid %in% .$multicancer_uid)) %>%
    arrange(multicancer_uid)
age2 <- bind_rows(ras, kras, nras)
write_csv(age2, file.path(tabdir, "age.csv"))
```


## Race

```{r race_data, fig.width=12, fig.height=7}
facet_labels <- function(uid){
    uid2 <- strsplit(uid, "~")
    caps <- Hmisc::capitalize
    cancer <- sapply(uid2, "[", 1) %>%
        str_replace_all("cancer of unknown primary", "CUP") %>%
        str_replace_all("non-small cell lung cancer", "NSCLC") %>%
        str_replace_all("pancreatic cancer", "PAC") %>%
        str_replace_all("ovarian cancer", "OVCA") %>%
        str_replace_all("colorectal cancer", "CRC") %>%
        str_replace_all("b-lymphoblastic leukemia/lymphoma",
                        "B ALL") %>%
        caps()
    gene <- sapply(uid2, "[", 2)
    codons <- sapply(uid2, "[", 3)
    flabels <- paste0(cancer, ", ", gene, " (", codons, ")") %>%
        setNames(uid)
    flabels
}

varlevels <- c("Multi-cancer", levels(race$variable))
mc <- multicancer %>%
    mutate(variable="Multi-cancer",
           variable=factor(variable, varlevels))
race2 <- race %>%
    select(-uid) %>%
    mutate(variable=factor(variable, varlevels)) %>%
    bind_rows(filter(mc, multicancer_uid %in% .$multicancer_uid)) %>%
    arrange(multicancer_uid)
write_csv(race2, file.path(tabdir, "race.csv"))
racelabels <- facet_labels(unique(race2$multicancer_uid))
race2 %>%
    mutate(variable=droplevels(variable),
           variable=fct_rev(variable)) %>%
    ggplot(aes(median, variable)) +
    geom_errorbarh(aes(xmin=q1, xmax=q2), height=0.1) +
    geom_point(pch=21, fill="white") +
    ylab("") +
    geom_text(x=Inf, aes(label=paste0(y, "/", n)),
              hjust=1, size=3) +
    theme_bw(base_size=15) +
    scale_x_continuous(expand=expansion(mult=0.5)) +
    theme(panel.grid=element_blank()) +
    xlab("Prevalence") +
    facet_wrap(~multicancer_uid, scales="free_x",
               labeller=labeller(multicancer_uid=racelabels)) +
    theme(strip.background=element_blank()) +
    geom_vline(data=filter(race2, variable=="Multi-cancer"),
               aes(xintercept=median), linetype="dashed",
               color="gray")
```


## Sex

```{r sex}
varlevels <- c("Multi-cancer", levels(sex$variable))
mc <- multicancer %>%
    mutate(variable="Multi-cancer",
           variable=factor(variable, varlevels))
sex2 <- sex %>%
    select(-uid) %>%
    mutate(variable=factor(variable, varlevels)) %>%
    bind_rows(filter(mc, multicancer_uid %in% .$multicancer_uid)) %>%
    arrange(multicancer_uid)
write_csv(sex2, file.path(tabdir, "sex.csv"))
```
