---
title: "RAS prevalence for specific amino acid substitutions"
site: workflowr::wflow_site
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
editor_options:
  chunk_output_type: console
---


```{r caching, echo=FALSE}
knitr::opts_chunk$set(autodep = TRUE)
```

```{r packages, message=FALSE, warning=FALSE}
library(IRanges)
library(here)
library(magrittr)
library(tidyverse)
library(ggplot2)
library(genie.6.1)
library(gridExtra)
library(kableExtra)
tabledir <- here("public", "table", "ras_amino_acid_prev.Rmd")
fs::dir_create(tabledir)
library(purrr)
library(viridis)
library(hrbrthemes)
##ipsum_pdf()
```

# Figures


```{r commandargs}
models <- here("output", "prevalence", "summarize_models.R",
               "models_aa.rds") %>%
    readRDS()
multicancer <- models %>%
    filter(cancer_group=="multi") %>%
    unnest("data") %>%
    mutate(Y=prettyNum(y, big.mark=","),
           N=prettyNum(n, big.mark=",")) %>%
    unite(nlabel, c(Y, N), sep="/", remove=FALSE) %>%
    group_by(hugo_symbol) %>%
    nest()
```


In the figures below, the grey point indicates the empirically estimated proportion and the white circle is the posterior median.

```{r prevalence, fig.align='center', fig.width = 20, fig.height = 25}
gg_ras <- function(hugo_symbol, data){
    x <- data %>%
        mutate(codon=codon_range)
    ## this is just to decide the order of the cancers
    xx <- filter(x, codon=="12C") %>%
        arrange(median)
    cancer_ <- xx$cancer
    x %>%
        mutate(cancer=factor(cancer, cancer_)) %>%
        filter(n > 150) %>%
        ggplot(aes(median, cancer)) +
        geom_errorbarh(height=0.2,
                       aes(xmin=q1, xmax=q2)) +
        geom_point(aes(x=y/n), color="gray50",
                   fill="gray70",
                   size=1, shape=21) +
        geom_point(color="steelblue", shape=21,
                   fill="white", size=1) +
        theme_bw(base_size=15) +
        theme(panel.grid=element_blank(),
              strip.background=element_blank()) +
        geom_text(aes(x=0.45, label=nlabel),
                  hjust=1, size=3) +
        ylab("") + xlab("") +
        facet_wrap(~codon) +
        ggtitle(hugo_symbol)
}
fig.list <- multicancer %>%
    pmap(gg_ras)
fig.list
```

# Stratified analyses

## KRAS 12C

For cancers with at least 100 individuals and an empirical RAS codon prevalence of at 10%, plot the stratified prevalence estimates.

```{r gg_strata}
gg_strata <- function(cancer, data){
    x <- data
    overall <- filter(x, variable=="Overall") %>%
        pull(median)
    x %>%
        mutate(variable=fct_rev(variable)) %>%
        filter(variable!="NA") %>%
        ggplot(aes(median, variable)) +
        geom_errorbarh(aes(xmin=q1, xmax=q2), height=0.2) +
        geom_point(pch=21, fill="white") +
        theme_bw(base_size=15) +
        facet_grid(confounder~., space="free_y",
                   scales="free_y") +
        geom_text(x=1, aes(label=paste0(y, "/", n)), hjust=1) +
        theme(panel.grid=element_blank(),
              strip.text.y=element_text(hjust=0, angle=0),
              strip.background=element_blank()) +
        geom_vline(xintercept=overall, color="gray",
                   linetype="dashed") +
        ylab("") +
        xlab("Prevalence") +
        xlim(c(0,1)) +
        ggtitle(cancer)
}
caps <- Hmisc::capitalize
## only makes sense to assess confounding for cancers with appreciable RAS variants
focus <- unnest(multicancer, data) %>%
    filter(y/n > 0.1, n > 100) %>%
    pull(cancer) %>%
    unique()
```

## KRAS 12C

```{r kras_12c,  fig.width=8, fig.height=7}
rasv <- "KRAS_12C"
models2 <- models %>%
    unnest(data) %>%
    filter(ras==rasv) %>%
    group_by(cancer) %>%
    nest()
fig.list <- models2 %>%
    filter(cancer %in% focus) %>%
    pmap(gg_strata)
fig.list
```

## KRAS 12D

```{r kras_12d,  fig.width=8, fig.height=7}
rasv <- "KRAS_12D"
models2 <- models %>%
    unnest(data) %>%
    filter(ras==rasv) %>%
    group_by(cancer) %>%
    nest()
fig.list <- models2 %>%
    filter(cancer %in% focus) %>%
    pmap(gg_strata)
fig.list
```

## KRAS 12V
```{r kras_12v,  fig.width=8, fig.height=7}
rasv <- "KRAS_12V"
models2 <- models %>%
    unnest(data) %>%
    filter(ras==rasv) %>%
    group_by(cancer) %>%
    nest()
fig.list <- models2 %>%
    filter(cancer %in% focus) %>%
    pmap(gg_strata)
fig.list
```

## NRAS 61R

```{r nras_61r,  fig.width=8, fig.height=7}
rasv <- "NRAS_61R"
models2 <- models %>%
    unnest(data) %>%
    filter(ras==rasv) %>%
    group_by(cancer) %>%
    nest()
fig.list <- models2 %>%
    filter(cancer %in% focus) %>%
    pmap(gg_strata)
fig.list
```


# Combined table

```{r combined_table, results="hide"}
dat <- models %>%
    ungroup() %>%
    unnest("data") %>%
    group_by(cancer_group, confounder)
dat.list <- dat %>%
    group_split()
unix_name <- function(x){
    x <- as.character(x) %>%
        tolower() %>%
        str_replace_all(" ", "_")
    x
}
grps <- dat %>%
    group_keys() %>%
    mutate(confounder=unix_name(confounder),
           confounder=ifelse(confounder=="_", "multicancer",
                             confounder),
           fname=paste0(confounder, ".csv"))
fnames <- here("public", "table", "ras_amino_acid_prev.Rmd", grps$fname)
dat.list %>%
    map2(fnames, write_csv)
##all_models <- model.list[[1]] %>%
##multicancer <- unnest(multicancer2, "data") %>%
##    select(-c(Y, N, nlabel)) %>%
##    rename(`2.5%`=q1,
##           `97.5%`=q2)
##write_csv(multicancer, here("public", "table", "ras.Rmd", "multicancer.csv"))
##dat4 <- unnest(dat3, "data") %>%
##    select(-overall) %>%
##    rename(`2.5%`=q1,
##           `97.5%`=q2)
##write_csv(dat4, here("public", "table", "ras.Rmd", "confounding.csv"))
```

# csv files

See [downloads](downloads.html) for results as plain-text csv files.
